---
title: "Chill portions and yield"
author: Katja Schiffers^a^, Cory Whitney^a^, Eduardo Fernandez^a^ and Eike Luedeling^a^ <br /> <small>^a^INRES-Horticultural Sciences, University of Bonn, Auf dem Huegel 6, 53121 Bonn, Germany <small/>
output:
    html_document:
bibliography: packages.bib
---
<!-- Spelling -->
<!-- The ABC √ option (upper right on the Rmarkdown console)-->

<!-- Grammar -->
<!-- devtools::install_github("ropenscilabs/gramr") -->
<!-- run_grammar_checker("HighDimensionalData.rmd”) -->

<!-- Print pdf and word versions -->
<!-- rmarkdown::render("Chill_Yield.Rmd", output_format = "pdf_document") -->
<!-- rmarkdown::render("Chill_Yield.Rmd", output_format = "word_document") -->


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
#packages in alphabetic order
library(bayesplot)
library(chillR)
library(devtools)
library(dplyr)
library(ggplot2)
library(ggpubr)
library(ggridges)
library(ggstance)
library(graphics)
#devtools::install_github("hortibonn/pasitR")
library(pasitR)
library(plyr)
library(reshape)
library(tidyverse) #tidyverse includes a number of useful packages
```

```{r, warning=FALSE, include = FALSE}
#Automatically write R package citation entries to a .bib file
knitr::write_bib(c(.packages(), 
                   'bayesplot', 
                   'chillR',
                   'decisionSupport',
                   'dplyr',
                   'ethnobotanyR',
                   'ggplot2', 
                   'ggridges', 
                   'graphics',
                   'plyr', 
                   'reshape', 
                   'tidyverse'), 'packages.bib')
```

# NOTES

Often data is limited 

1. Lack of data
1. How can we go about making assessments when we hve limited data?
1. when attempting to assess the relationships between temperature and yield
1. perhaps this data limitation does not need to hinder our assessment
1. Maybe a lot of data is not necessary
1. embrace uncertainty
1. preciseness is pointless
1. far into the future we have no data. We necesarily have uncertainty in our model forecasts

We use the pasitR package [@R-pasitR] in the R programming language [@R-base].

Create random chill and yield data using `sample()`. 

```{r}
chill <- sample(x = 1:50, size = 20, replace = TRUE)
yield <- sample(x = 1000:5000, size = 20, replace = TRUE)
```

Using `chillscatter()` create a chill portions (x) and yield (y) scatter plot with associated and estimated densities with loess smooth linear fits density curves. Plot made with the `scatter.hist()` function in the `plyr()` package [@R-plyr].  

```{r, fig.width=4, fig.height=4}
pasitR::chillscatter(chill, yield)
```

Use `chillkernel()` or another type of graphical representation of the same relationship. `chillkernel()` performs a two-dimensional kernel density estimation for yield and chill using the `kde2d()` function in the `MASS()` package [@R-MASS]. The function produces a matrix of the estimated density (z) of yield (y) and chill (x). As the density function restricts the shape of the kernel to a bivariate normal kernel, it looks slightly different compared to the scatter plot estimates above.

```{r, fig.width=4, fig.height=4}
pasitR::chillkernel(chill, yield)
```

`chillkernel()` shows a density surface plot of chill portions (x) and yield (y). The legend shows the value for the estimated density (z). The plot is made with the `filled.contour()` function of the `graphics()` package [@R-graphics].

In `chillkernel()` the density (z) over the entire plot integrates to one, and therefore represents the relative probability of an observation (yield along y-axis) given a specific chill portion (along x-axis). 

## Estimated yield given the expected chill

`chillkernelslice()` calculates the estimated yield given the expected chill, based on a slice of 'z' from the Kernel density calcualted with `chillkernel()`. The `expectedchill` parameter is set to 30.

```{r, fig.width=4, fig.height=4}
pasitR::chillkernelslice(chill, yield, expectedchill = 30)
```

`chillkernelslice()` plots the probabilities (shown along the y-axis) for the expected yield (shown along the x-axis). Since this is a cut through the density kernel `chillkernel()`, which integrates to 1, the probability values are relative, not absolute measures.

## Chill portion intervals

`chillviolin()` determines different possible chill portion intervals by calculating the optimal interval width for chill portions using the `IQR()` function in the `stats()` package, after the Freedman-Diaconis rule (IQR = interquartile range) [@R-base].
`Optimal interval width for our sample = 2 * interquartile range for our sample / (total number of observations for the interquartile range for our sample)^(1/3)`

```{r, fig.width=10, fig.height=4}
pasitR::chillviolin(chill, yield)
```

`chillviolin()` shows violin plots of chill portions (x) and yield (y) with six different intervals of chill portions. Plot made with `ggplot2()` [@R-ggplot2].

## Probability of yield given chill

`chillkernelslicerange()` The chill portion intervals, optimized interquartile ranges shown in `chillviolin()` can be used to select a range to slice from the density kernel `chillkernel()` as was doen for a single chill value in `chillkernelslice()`. Here we set the maximum chill to 20 and the minimum to 10.

```{r, fig.width=4, fig.height=4}
pasitR::chillkernelslicerange(chill, yield, 10, 20)
```

`chillkernelslicerange()` plots the probabilities (shown along the y-axis) for the expected yield (shown along the x-axis). As with `chillkernelslice()` the probability values shown are relative, not absolute measures. They are the result of cuts through the density kernel `chillkernel()`, which integrates to 1. 

# section

The functions here closely follow chillR [@R-chillR] and decisionSupport [@R-decisionSupport].

## References
